---
title: "HW 1"
subtitle: "Data 180, Professor Kennedy"
author: 
<<<<<<< HEAD
  name: "Melanie Duong"
  email: "duongng@dickinson.edu"
=======
  name: "melanie duong"
  email: "XXXX@dickinson.edu"
>>>>>>> main
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document
---
<<<<<<< HEAD
Make sure that you have accepted the hw invitation link, and working with your clone repository.
=======
Make sure that you have accepted the hw invitation link, and working with your clone repository
>>>>>>> main

**Due date:** September 21 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

## Preliminary Task
Go to [datacamp.com](https://datacamp.com) and complete the tutorial "Introduction to R".


## Task 1
a) Store the data below in a data frame object named `my_df`.


|Person | Preference|
|:------|:----------|
|Alice | Twitter |
| Bob | Facebook |
| Charlie | Twitter |
| Denise | Instagram |
| Erin | Facebook |
| Frank | Twitter |
| Gisele | Twitter |
| Harry | Instagram |

<<<<<<< HEAD
```{r}
my_df <- data.frame(
  Person = c("Alice", "Bob", "Charlie", "Denise", "Erin", "Frank", "Gisele", "Harry"),
  Preference = c("Twitter", "Facebook", "Twitter", "Instagram", "Facebook", "Twitter", "Twitter", "Instagram")
)
my_df
```


b) What are the dimensions of `my_df`?

```{r}
dim(my_df)
print("The dimensions of my_df are 8 rows and 2 columns.")
```


c) Write R code that counts the total number of users.

```{r}
total_users <- nrow(my_df)
total_users
print("Total number of users is 8.")
```

d) Write R code that counts the total number of users who prefer Twitter.

```{r}
total_twitteruser <- sum(my_df$Preference == "Twitter")
total_twitteruser
print("Total number of users who prefer Twitter is 4.")
```

e) Write R code that returns what Gisele prefers.

```{r}
giselePrefer <- my_df$Preference[my_df$Person == "Gisele"]
giselePrefer
print("Gisele prefers Twitter.")
```

## Task 2
a) Create a vector named `my_vec` that includes all integers from 1 to 10 (inclusive).

```{r}
my_vec <- seq(1, 10)
my_vec
```

b) Replace the second element in `my_vec` with the number `99`.

```{r}
my_vec[2] <- 99
my_vec
```

c) Write a function called `my_func` that prints `This element is greater than 5!` if an element in `myvec` is greater than 5, and `This element is NOT greater than 5!` otherwise.

```{r}
my_func <- function(my_vec){
  for (num in my_vec) {
    if (num > 5){
      print("This element is greater than 5!")
    } else {
      print("This element is NOT greater than 5!")
    }
  }
}
my_func(my_vec)
```

## Task 3
a. Write a function named `get_fibonacci` that returns the n'th value in the [fibonacci sequence](https://www.mathsisfun.com/numbers/fibonacci-sequence.html). For example, `get_fibonacci(5)` should return 3; `get_fibonacci(7)` should return 8 and so on. ***Hint:** It is possible to call the same function from inside it. E.g., you can call `get_fibonacci()` from inside `get_fibonacci()`. Like inception!*

```{r}
get_fibonacci <- function(n) {
  if (n <= 0) {
    return("Please enter a positive integer")
  } else if (n == 1) {
    return(0)
  } else if (n == 2) {
    return(1)
  } else {
    return(get_fibonacci(n - 1) + get_fibonacci(n - 2))
  }
}
```

b. Using your function in part (a), find and return the 19th value in the fibonacci sequence.

```{r}
n <- 19
get_fibonacci(n)
```
=======
b) What are the dimensions of `my_df`?

c) Write R code that counts the total number of users.

d) Write R code that counts the total number of users who prefer Twitter.

e) Write R code that returns what Gisele prefers.

## Task 2
a) Create a vector named `my_vec` that includes all integers from 1 to 10 (inclusive).

b) Replace the second element in `my_vec` with the number `99`.

c) Write a function called `my_func` that prints `This element is greater than 5!` if an element in `myvec` is greater than 5, and `This element is NOT greater than 5!` otherwise.

## Task 3
a. Write a function named `get_fibonacci` that returns the n'th value in the [fibonacci sequence](https://www.mathsisfun.com/numbers/fibonacci-sequence.html). For example, `get_fibonacci(5)` should return 3; `get_fibonacci(7)` should return 8 and so on. ***Hint:** It is possible to call the same function from inside it. E.g., you can call `get_fibonacci()` from inside `get_fibonacci()`. Like inception!*

b. Using your function in part (a), find and return the 19th value in the fibonacci sequence.

>>>>>>> main

<br>
You are done! 🏁 Don't forget to `commit` and `push` your .Rmd file to your Github clone repository before the due date.
<br>

